---
title: "ALFAM2 confidence interval calculations"
author: "Sasha D. Hafner"
date: "`r Sys.Date()`"
---

# Overview
In January/February 2024 I added some code for calculating confidence intervals to `alfam2()`, given multiple parameter sets that themselves represent the distribution of possible parameter values.
This document demonstrates the implementation.

# ALFAM2 package

This is from the dev branch.
Use line below to remove existing version and install latest dev version.

```{r}
#remove.packages('ALFAM2') ; devtools::install_github('sashahafner/ALFAM2', ref = 'dev')
```

```{r}
library(ALFAM2)
packageVersion('ALFAM2')
```

I am using v3.70 here.

# Example 1 with dynamics

```{r}
dat <- data.frame(ctime = 0:84*2, TAN.app = 100, man.dm = 8, 
                  air.temp = 7 + 7*sin(0:84*2 * 2*pi/24) + rnorm(85, 0, 2), 
                  wind = 10^(0.5 + 0.4*sin(0:84*2 * 2*pi/24) + 
                             rnorm(85, 0, 0.12)), 
                  app.mthd = "bc")
```

```{r}
plot(air.temp ~ ctime, data = dat, type = 'o', col = 'gray45')
plot(wind ~ ctime, data = dat, type = 'o', col = 'blue')
```

Normal call without confidence intervals (CI).

```{r}
pred1 <- alfam2(dat, app.name = 'TAN.app', time.name = 'ctime', warn = FALSE)
head(pred1)
```

Add CI.
They are given in the output with `.lwr` and `.upr` suffixes.

```{r}
predci <- alfam2(dat, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', warn = FALSE, conf.int = 0.90, 
                pars.ci = alfam2pars03var_alpha)
head(predci)
```

By default CI are only returned for variable `er` = relative cumulative emission.

```{r}
plot(er ~ ctime, data = predci, type = 'l', ylim = c(0, max(predci$er.upr)))
lines(er.lwr ~ ctime, data = predci, type = 'l', col = 'blue')
lines(er.upr ~ ctime, data = predci, type = 'l', col = 'red')
```

This 90% CI is quite wide, it is seems a bit strange that the prediction with default parmeters is close to the upper limit at the start.
But these are draft parameter values, so this second issue may not be present in the final values.
The CIs will likely remain wide.

We can add any output variables for CI calculation, but `quantile` is applied by variable, so the limits may not associate between variables.
Use the `var.ci` argument for this.
Here we request 3 variables.

```{r}
predci <- alfam2(dat, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', warn = FALSE, conf.int = 0.90, 
                pars.ci = alfam2pars03var_alpha, var.ci = c('er', 'j', 'r1'))
head(predci)
```

Note that times with any `NaN` etc. in one of `var.ci` columns will be dropped before applying the `quantile()` function.
So here all `lwr` and `upr` limits are `NA` for time = 0 h.

```{r}
plot(er ~ ctime, data = predci, type = 'l', ylim = c(0, max(na.omit(predci$er.upr))))
lines(er.lwr ~ ctime, data = predci, type = 'l', col = 'blue')
lines(er.upr ~ ctime, data = predci, type = 'l', col = 'red')
```

```{r}
plot(j ~ ctime, data = predci, type = 's', ylim = c(0, max(na.omit(predci$j.upr))))
lines(j.lwr ~ ctime, data = predci, type = 's', col = 'blue')
lines(j.upr ~ ctime, data = predci, type = 's', col = 'red')
```

# Example 2 application to multiple groups

Here is a test where each group has a different incorporation time (based on one in vignette).

```{r}
datm <- data.frame(scenario = 1:6, ctime = 168, TAN.app = 50, 
                   man.dm = 8, air.temp = 20, wind.2m = 4, 
                   app.mthd = 'bc',
                   incorp = 'deep', 
                   t.incorp = c(0.1, 1, 6, 24, 168, NA))
```


```{r}
predci <- alfam2(datm, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', time.incorp = 't.incorp', group = 'scenario',
                conf.int = 0.90, 
                pars.ci = alfam2pars03var_alpha, var.ci = c('er'))
predci
```

Plot emission versus time of incorporation.

```{r}
plot(datm$t.incorp, predci$er, type = 'o', ylim = c(0, max(predci$er.upr)))
lines(datm$t.incorp, predci$er.lwr, type = 'o', col = 'blue')
lines(datm$t.incorp, predci$er.upr, type = 'o', col = 'red')
```

Notice how the CI is larger for rapid incorporation, because of uncertainty in incorporation parameters that has a smaller effect when incorporation is done later.

By default the model is run with all the different parameter sets provided in `pars.ci`, which is 100 in this draft object.

```{r}
dim(alfam2pars03var_alpha)
```

For speed some users might want to sometimes reduce that.

```{r}
predci <- alfam2(datm, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', time.incorp = 't.incorp', group = 'scenario',
                conf.int = 0.90, pars.ci = alfam2pars03var_alpha,
                n.ci = 10)
predci
```

# Example 3 on getting all predictions
When `conf.int=` some number it is used in the `quantile` function.
To get all results, use `conf.int = 'all'`.
This could be useful to combine uncertainty in parameter values with uncertainty in inputs.
Of course then the CIs would have to be constructed outside the `alfam2` function, but it cannot do everything, so I think this is OK.

```{r}
datvar <- data.frame(ctime = 168, TAN.app = 50, man.dm = rnorm(7, mean = 8, sd = 1), 
                     air.temp = 20, wind.2m = 3)
```

Here is what we get with numeric `conf.int`.

```{r}
predci <- alfam2(datvar, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', group = 'man.dm',
                conf.int = 0.90, pars.ci = alfam2pars03var_alpha)
predci
```

Not so useful.

The alternative:


```{r}
predci <- alfam2(datvar, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', group = 'man.dm',
                conf.int = 'all', pars.ci = alfam2pars03var_alpha, n.ci = 3)
predci
```

21 rows here, deliberately small so we can look at the results.

More plausible usage would have at least 100 of each I suppose, for 10000 rows in the output.

```{r}
datvar <- data.frame(ctime = 168, TAN.app = 50, man.dm = rnorm(100, mean = 8, sd = 1), 
                     air.temp = 20, wind.2m = 3)
```

```{r}
predci <- alfam2(datvar, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', group = 'man.dm',
                conf.int = 'all', pars.ci = alfam2pars03var_alpha, n.ci = 100)
head(predci)
```

And then, externally, for a 95% confidence interval that includes uncertainty in both inputs (only DM here) and parameters, we can use `quatile()`:

```{r}
quantile(predci$er, c(0.05, 0.95))
```

# Error messages

The calls below demonstrate some errors.

```{r,error=TRUE}
predci <- alfam2(datvar, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', group = 'man.dm',
                conf.int = 'all', n.ci = 100)
```

```{r,error=TRUE}
predci <- alfam2(datvar, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', group = 'man.dm',
                conf.int = 'all', pars.ci = alfam2pars03var_alpha, n.ci = 1000)
```

```{r,error=TRUE}
predci <- alfam2(datvar, pars = alfam2pars03_alpha, app.name = 'TAN.app', 
                time.name = 'ctime', group = 'man.dm',
                conf.int = 'all', pars.ci = alfam2pars03var_alpha, n.ci = 10, 
                var.ci = 'blahblah')
```


