---
title: 'Repeat evaluation of ALFAM2 parameter set 1'
author: 'Sasha D. Hafner'
date: '28 January 2024'
urlcolor: blue
---

# Overview
This document for Arman Favrot presents a repeat of the evaluation of the ALFAM2 model with parameter set 1 originally presented in the 2019 paper introducing the model.

# Packages and functions

```{r}
library(data.table)
library(ALFAM2)
source('dfsumm.R')
source('model_stats.R')
```

See ALFAM2 version.

```{r}
packageVersion('ALFAM2')
```

The problem you found with duplicated columns should not be present in this version, which is available from the dev branch now.
But note that the function does not add dummy variable columns for "reference levels", and both cattle manure and trailing hose are reference levels, so that behavior was not an error.

# Input data

Get pmid from paper.

```{r}
pmidcal <- fread('../data/S1_plot_codes_calibration.csv')
pmideval <- fread('../data/S2_plot_codes_evaluation.csv')
```

To associate these with measurements below, combine them in a data table here.

```{r}
pmidcal[, datasub := 'cal']
pmideval[, datasub := 'eval']
pmid <- rbind(pmidcal, pmideval)
```

Load interval-level data.
Note this is version 1.0 of ALFAM2 "database".
See <https://github.com/sashahafner/ALFAM2-data/tree/cfa0055e44907578bf40e0eac389e020f93dd6b1/data%20-%20ALFAM2%20output> for this version.
Unfortunately I was just getting started with GitHub releases around this time in 2018, so there is no releaase for this version.
Since then the situation has improved.
Now you can always find the database version in `data-output/...` ([here](https://github.com/sashahafner/ALFAM2-data/tree/master/data-output/03)), and new versions get releases ([here](https://github.com/sashahafner/ALFAM2-data/releases)).

```{r}
idat <- fread('../data/ALFAM2_interval.csv')
```

Fix changed variable name.

```{r}
idat[, app.mthd := app.method]
```

Merge in pmid and subset keys.

```{r}
dim(idat)
idat <- merge(idat, pmid, by = 'pmid')
dim(idat)
```

Trim to 78 hours.

```{r}
idat <- idat[ct > 0 & ct < 78, ]
```

Check values.

```{r}
dfsumm(idat[datasub == 'eval',.(pmid, app.mthd, app.rate, man.dm, man.source, air.temp, wind.2m, 
                                man.ph, rain.rate, incorp)])
dfsumm(idat[datasub == 'cal',.(pmid, app.mthd, app.rate, man.dm, man.source, air.temp, wind.2m, 
                               man.ph, rain.rate, incorp)])
```

Fill in missing wind speed, air temperature, rainfall values.
This is a thorny problem. 
Dropping values means leaving out emission intervals.
Here they are set to the average by plot, or, for rain, zero.

```{r}
idat[, wind.2m.ave := mean(wind.2m, na.omit = TRUE), by = pmid]
idat[, air.temp.ave := mean(air.temp, na.omit = TRUE), by = pmid]

idat[is.na(wind.2m), wind.2m := wind.2m.ave]
idat[is.na(air.temp), air.temp := air.temp.ave]

idat[is.na(rain.rate), rain.rate := 0]
idat[is.na(rain.cum), rain.cum := 0]
```

Check values.

```{r}
dfsumm(idat[datasub == 'eval',.(pmid, app.mthd, app.rate, man.dm, man.source, air.temp, wind.2m, 
                                man.ph, rain.rate, incorp)])
dfsumm(idat[datasub == 'cal',.(pmid, app.mthd, app.rate, man.dm, man.source, air.temp, wind.2m, 
                               man.ph, rain.rate, incorp)])
```

# Model application
Get parameters *without pH*.
The pH parameters only apply to acidified slurry in set 1.
This has changed in set 2 and the new set 3.

```{r}
pars <- alfam2pars01[!grepl('man.ph', names(alfam2pars01))]
```

```{r}
pars
```

Generate predictions.
Note `group` argument allows application to multiple plots.
See the vignette for more details.
So you can apply the function to any number of plots/locations with a single call.
And in this version of the package `prep.dum = TRUE` by default (new argument name too).

```{r}
pred <- alfam2(idat, pars = pars, app.name = 'tan.app', time.incorp = 'time.incorp', group = 'pmid')
```

```{r}
head(pred)
```

See the dummy variables in the output.

Add predictions to the input variable data frame.

```{r}
idat[, `:=` (j.NH3.pred = pred$j, e.cum.pred = pred$e, e.rel.pred = pred$er)]
```

And get final values (not beyond 78 hours).

```{r}

idat[, ct.max := max(ct), by = pmid]
idat.final <- idat[ct == ct.max, ]
```

# Model fit

Flux.

```{r}
summ1 <- idat[, .(n = length(j.NH3), 
                  me = me(j.NH3, j.NH3.pred),
                  mae = mae(j.NH3, j.NH3.pred),
                  mbe = mbe(j.NH3, j.NH3.pred)
                  ), by = datasub]
summ1
```

Cumulative emission.

```{r}
summ2 <- idat.final[, .(n = length(e.cum), 
                        me.cum = me(e.cum, e.cum.pred), me.rel = me(e.rel, e.rel.pred),
                        mae.cum = mae(e.cum, e.cum.pred), mae.rel = mae(e.rel, e.rel.pred),
                        mbe.cum = mbe(e.cum, e.cum.pred), mbe.rel = mbe(e.rel, e.rel.pred)
                        ), by = datasub]
summ2
```

summ



