---
title: 'ALFAM2 model fit with parameter set 3'
output: pdf_document
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y %H:%M')`"
---

# 1. Summary
Determination of ALFAM2 model performance using parameter set 3, for the 700+ plots used for parameter estimation.

# 2. Setup

Some functions.

```{r}
source('functions/interpm.R')
source('functions/dfsumm.R')
source('functions/model_stats.R')
```

Packages.

```{r}
library(data.table)
library(ggplot2)
```

ALFAM2, check version.

```{r}
library(ALFAM2)
packageVersion('ALFAM2')
alfam2pars03
```

And it is good practice to check hash, in case someone (me) did a poor job with version numbers.

```{r}
devtools::package_info('ALFAM2')
```

# 3. Measurement data

See `data-emission/data/data_version.txt` for version.
This was copied directly from the 20204 paper (<https://github.com/aU-BCE-EE/hafner-2024-ALFAM2-dev>) so is already outdated.
Change `rtag` in `data-emission/scripts/main.R` to select a different version.

```{r}
idat <- fread('data-emission/data/ALFAM2_interval.csv')
pdat <- fread('data-emission/data/ALFAM2_plot.csv')
```

The `pmid` keys from the 2024 paper (<https://github.com/aU-BCE-EE/hafner-2024-ALFAM2-dev>).

```{r}
pmid1 <- fread('pmid/pmid_sub1.txt')[[1]]
```

Subset to subset 1 from 2024 paper.

```{r}
dim(pdat)
pdat <- pdat[pmid %in% pmid1, ]
dim(pdat)
```

Merge into idat

```{r}
dim(idat)
idat1 <- merge(idat, pdat, by = c('pid', 'pmid'))
dim(idat1)
```

New names.

```{r}
idat1[, wind.sqrt := sqrt(wind.2m)]
idat1[, app.mthd := app.method]
```

Check for missing values.

```{r}
dfsumm(idat1[, .(app.mthd, man.dm, man.source, man.ph, tan.app)])
dfsumm(idat1[, .(ct, cta, air.temp, wind.sqrt, rain.rate, rain.cum)])
```

Plenty, as in paper.

Fill in missing pH with institute means from full database

```{r}
mnph <- pdat[, .(man.ph.mean = mean(na.omit(man.ph)), man.ph.n = sum(is.na(man.ph))), by = inst]
mnph
idat1 <- merge(idat1, mnph, by = 'inst')
idat1[is.na(man.ph), man.ph.missing := TRUE]
idat1[is.na(man.ph), man.ph := man.ph.mean]
```

Interpolate missing wind and air temperature values.
Set missing rain to 0.
Set `cta` to `ct` where missing.
And drop obs with `cta < 0`.

```{r}
idat1[, `:=` (interp.wind = is.na(wind.2m), interp.air.temp = is.na(air.temp)), ]
idat1 <- interpm(idat1, 'ct', c('wind.sqrt', 'air.temp'), by = 'pmid', rule = 2)

idat1[is.na(rain.rate), rain.missing := TRUE]
idat1[is.na(rain.rate), rain.rate := 0]
idat1[is.na(rain.cum), rain.cum := 0]

idat1[is.na(cta), cta := ct]

idat1 <- idat1[cta > 0, ]
```

Check for other missing values.

```{r}
dfsumm(idat1[, .(app.mthd, man.dm, man.source, man.ph, tan.app)])
dfsumm(idat1[, .(ct, cta, air.temp, wind.sqrt, rain.rate, rain.cum)])
```

Everything is OK now.

# 4. ALFAM2 predictions

```{r}
args(alfam2)
pred <- alfam2(idat1, pars = alfam2pars03, app.name = 'tan.app', time.name = 'cta',
	       group = 'pmid', 
	       time.incorp = 'time.incorp')
```

Combine with measurements.

```{r}
out <- merge(idat1, pred, by = c('pmid', 'cta'), suffixes = c('', '.pred'))
```

Error.

* in cumulative emission, kg N / ha,
* in relative emission, fraction of applied TAN,
* and relative to measured cumulative emission, as a percentage

```{r}
out[, err.e.cum := e - e.cum]
out[, err.er := e.rel - er]
out[, err.rel := 100 * err.e.cum / e.cum]
hist(out$err.e.cum)
hist(out$err.er)
```

Get latest time.

```{r}
out[, cta.max := max(cta), by = .(pmid)]
outfinal <- out[cta == cta.max, ]
```

Take a look.

```{r}
ggplot(out, aes(er, e.rel, colour = app.mthd)) + geom_point()
ggplot(outfinal, aes(er, e.rel, colour = app.mthd)) + geom_point()
```

## Fit statistics.

Absolute emission kg N / ha

```{r}
rmse(m = outfinal$e.cum, p = outfinal$e)
mae(m = outfinal$e.cum, p = outfinal$e)
me(m = outfinal$e.cum, p = outfinal$e)
```

Relative emission, fraction applied TAN

```{r}
rmse(m = outfinal$e.rel, p = outfinal$er)
mae(m = outfinal$e.rel, p = outfinal$er)
me(m = outfinal$e.rel, p = outfinal$er)
```

Mean and median relative error

```{r}
mean(abs(outfinal$err.rel))
median(abs(outfinal$err.rel))
quantile(abs(outfinal$err.rel))
```

And by application method (low emission methods have high relative error).

```{r}
outfinal[, .(mn = mean(abs(err.rel)), md = median(abs(err.rel))), by = .(app.mthd)]
```

Also quite poor for broadcast `bc`!



